---
title: "DATA-413/613"
subtitle: "Things to Remember I"
author: "Hamid Semiyari"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 4
    number_sections: true
    theme: cerulean
urlcolor: "blue"
linkcolor: "green"
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.height = 3, 
                      fig.width  = 6,
                      fig.align  = "center")
ggplot2::theme_set(ggplot2::theme_bw())
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

-   A data that contains the following three attributes may consider as
    Big Data

    1.  Huge Volume: More and more data becomes rapidly available
    2.  High Velocity: Data is being generated in a very fast rate
    3.  Hulking Variety: Data comes from in many forms.

-   There are three types of (Big) data

    1.  Structured data: is search-able, organized in a predefined data
        model, Stored in Rows and Columns. Example: Social Security
        Number, Customer Names, geolocation, stock information, digital
        photographs, where image has no predefined structure but has
        certain structural attributes...
    2.  Semi-Structured data: Organized, but not very well. Example:
        Tweet organized by hashtags,
    3.  Unstructured data: Not easy to search, Not predefined data
        model, Stored in various format. Example: Emails, Voice message
        files, image files,..

-   What is API?\
    It is short for “Application Programming Interface.“ API is some
    kind of interface which has a set of functions that allow
    programmers to access specific features or data of an application,
    operating system or other services. Web API as the name suggests, is
    an API over the web which can be accessed using HTTP protocol. It is
    a concept and not a technology. We can build Web API using different
    technologies such as Java, .NET etc

-   **Tidying** your data means storing it in a consistent form that
    matches the semantics of the data-set with the way it is stored. In
    brief, when your data is tidy, each column is a variable, and each
    row is an observation. Tidy data is important because the consistent
    structure lets you focus your struggle on questions about the data,
    not fighting to get the data into the right form for different
    functions.

-   **Data Semantics** is the study of the meaning and use of specific
    pieces of data in computer programming and other areas that employ
    data. When studying a language, semantics refers to what individual
    words mean and what they mean when put together to form phrases or
    sentences. In data semantics, the focus is on how a data object
    represents a concept or object in the real world. It helps you
    discover valuable real-world relationships.

    -   In a banking system, various types of data hold different
        meanings and have specific structures. Here are some examples:

        1.  **Customer Information**: This includes data about bank
            customers such as their name, address, contact information,
            account number, and account balance.

        2.  **Transaction Data**: This consists of records of
            transactions made by customers, including details such as
            transaction type (deposit, withdrawal, transfer), amount,
            date, and time.

        when designing a banking application, developers need to define
        the structure and meaning of various data elements and ensure
        that operations performed on this data adhere to the intended
        semantics. This helps in maintaining data integrity,
        consistency, and reliability within the system.

-   **Transformation** includes narrowing in on observations of interest
    (like all people in one city, or all data from the last year),
    creating new variables that are functions of existing variables
    (like computing speed from distance and time), and calculating a set
    of summary statistics (like counts or means). Together, tidying and
    transforming are called **wrangling**, because getting your data in
    a form that’s natural to work with often feels like a fight!

-   **Hadoop** is an open source framework that is used to efficiently
    store and process large datasets ranging in size from gigabytes to
    petabytes2 of data.

-   **Spark** is also open source — is a data processing engine for big
    data sets. Like Hadoop, Spark splits up large tasks across different
    nodes. However, it tends to perform faster than Hadoop and it uses
    random access memory (RAM3) to cache 4 and process data instead of a
    file system. This enables Spark to handle use cases that Hadoop
    cannot.

-   Non-rectangular data.\
    we have seen rectangular data: collections of values that are each
    associated with a variable and an observation. There are lots of
    datasets that do not naturally fit in this paradigm, including
    images, sounds, trees, and text.
